---
title: "Run `NAVAR` experiments"
subtitle: "NAVAR"
author:
- "Marc Agustí (marc.agusti@barcelonagse.eu)"
- "Patrick Altmeyer (patrick.altmeyer@barcelonagse.eu)"
- "Ignacio Vidal-Quadras Costa (ignacio.vidalquadrascosta@barcelonagse.eu)"
date: "`r format(Sys.Date(), '%B, %Y')`"
output: 
  bookdown::html_document2: default
  github_document: default
  bookdown::pdf_document2: 
    toc: false
bibliography: "bib.bib"
link-citations: true
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(reticulate)
```

# Setup

In this file we just explore the NAVAR code in some more detail by first just running the experiments step-by-step to get a better idea of how the code works. The code chunk below just install requirements.

```{bash, eval=FALSE}
python3 -m pip install -r requirements.txt
```

# DREAM3 experiments

We start by importing the relevant modules for these experiments as in `run_dream_experiments.py`.

```{python}
from train_NAVAR import train_NAVAR
from evaluate import calculate_AUROC, dream_file_to_causal_matrix
import pandas as pd
import argparse
```

To understand the code better, we will just run a single experiment, in particular the one called `ecoli1`.

```{python}
experiment = 'ecoli1' # experiment name
# should accuracy be evaluated?
evaluate = True
# should we use LSTM?
lstm = True 
lambda1 = 0.26025947107502856
batch_size = 46 # batch size for training
wd = 1.4961159190152877e-05
hidden_nodes = 10
learning_rate = 0.002 
hl = 1 # number of hidden layers
```

Given the provided parameters we will import the data. Instead of the 5000 (!) epochs that @bussmann2020neural use, we will just use 5.

```{python}
# load the data
file = f'experiments/{experiment}.tsv'
ground_truth_file = f'experiments/{experiment}_gt.txt'
data = pd.read_csv(file, sep='\t')
data = data.values[:, 1:]
epochs = 1
maxlags = 21 if lstm else 2
```

Let's explore that data for a moment. The `ecoli1` data is made up by 966 rows and 100 columns.

```{r}
library(data.table)
dt <- data.table(py$data)
dim(dt)
```

## Running the experiment

Next we will turn to actually running the experiment as in `run_dream_experiments.py`. 

```{python}
import torch
import numpy as np
from dataloader import DataLoader
from NAVAR import NAVAR, NAVARLSTM
```


```{python}
# Other parameters
dropout=0
hidden_layers=hl

# T is the number of time steps, N the number of variables
T, N = data.shape

# initialize the NAVAR model
if lstm:
    model = NAVARLSTM(N, hidden_nodes, maxlags, dropout=dropout, hidden_layers=hidden_layers)
else:
    model = NAVAR(N, hidden_nodes, maxlags, dropout=dropout, hidden_layers=hidden_layers)
```




```{python}
# start training
print(f"Starting training on the data from experiment {experiment}, training for {epochs} iterations.")
score_matrix, _, _ = train_NAVAR(data, maxlags=maxlags, hidden_nodes=hidden_nodes, dropout=0, epochs=epochs,
                                 learning_rate=learning_rate, batch_size=batch_size, lambda1=lambda1,
                                 val_proportion=0.0, weight_decay=wd, check_every=1, hidden_layers=hl, normalize=True,
                                 split_timeseries=21, lstm=lstm)
# evaluate
print('Done training!')
if evaluate:
    ground_truth_matrix = dream_file_to_causal_matrix(ground_truth_file)
    AUROC = calculate_AUROC(score_matrix, ground_truth_matrix, ignore_self_links=True)
    print(f"The AUROC of this model on experiment {experiment} is: {AUROC}")
```




